\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-09-06 paper.tex
Submitted version?

\usepackage[draft,nomargin, marginclue, footnote]{fixme}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\columnwidth]{grounding_robot.pdf}
\caption{Robot reasoning about HRI and anticipation of human activities.
  Sources of information are multi-modal dialogue, and monitoring of the
  environment and human activity. The robot must adapt on-line its behaviours
  by merging computed plans with reactive control. The robot explicitly reasons
  on the fact that it is (or not) observed by the human. Reasoning and planning
  take place at symbolic as well as geometric level and systematically take
  into account both agents beliefs and activities (from the point of view of
  the robot).}

@@

\fxwarning{Need to carefully introduce this section: how did we come up this this architecture? Describe archi}

Over the last years, we have been slowly ramping up explicit knowledge
representation and manipulation in the deliberative and executive layers of our
robots. Ranging from situation assessment to symbolic task planning, from
verbal interaction to event-driven execution control, we have built up a
\emph{knowledge-oriented} architecture which is now used on a daily basis on
our robots.

Figure~\ref{fig|archi} gives an overview of the connections between the
deliberative components of our architecture~\cite{Alami2011}.

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may in return
trigger physical behaviours.

@@

HATP originality resides in its ability to produce plans for the robot's
actions \emph{as well as for the other participants} (humans or robots): the
robot plans for itself, but also for its interactors, which effectively endows the robot
with new interaction capabilities: It can partially predict the behaviour, and
thus monitor the engagement, of its human partner; the planner generates
synchronisation points (arrow going from the robot stream to the human stream on
Figure~\ref{plan_hatp1}) between the agents; the controller can also verbalize
the plans to explain its partner how a task may be shared~\cite{warnier2012when}.

@@

This allows us to write reactive robot controllers with a high level of
expressiveness: for instance, by subscribing to the event \stmt{human1 desires
?action, ?action type Look, ?action hasGoal myself}, we could trigger a
behaviour when the human expresses (through dialogue, gestures...) that he
wants to look at the robot itself.

@@

The second one involves the SHARY execution controller and the HATP symbolic
task planner besides ORO and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. Behaviours and motions
are fully planned.

@@

\fxwarning{Robot control related work -> what is the originality of SHARY? why is it relevant from an AI point of view? (Mathieu, Aurelie)}

Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture called
\textsc{CRAM} (Cognitive Robot Abstract Machine) that integrates
\textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
Prolog. Its underlying storage is based on an OWL ontology, derived from
\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KnowRob} have been demonstrated on
several real-world scenarios, where natural language recipes extracted from the
Internet had to be translated into plans and executed in a kitchen environment,
perceived and rebuilt on-line by the robots.

\fxwarning{robot control architecture for service robots: what else?}

@@

\fxwarning{What are our contribution regarding perspective taking? (Mathieu)}
\fxwarning{Mention other source of knowledge about humans -- including social rules for instance}

@@

\paragraph{Reasoning with the humans}

\fxwarning{Mention we represent the human knowledge as well (Severin)}

@@

\fxwarning{Introduce the paragraph}

@@

\fxwarning{Rewrite conclusion}

We have devised a decisional framework for human-robot interactive
task achievement that is aimed to allow the robot not only to
accomplish its tasks but also to produce behaviors that support its
engagement vis-a-vis its human partner and to interpret human
behaviors and intentions. 
Together and in coherence with this framework, we have developed
and experimented various task planners and interaction schemes that
allow the robot to select and perform its tasks while taking into
account explicitly the human abilities as well as the constraints
imposed by the presence of humans, their needs and preferences.

In this paper we have presented a decisional framework designed for
robots operating in a human environment. Our objective is to provide a
management of human-robot interaction that is an integral part of a
general robot control architecture.  This was done in order to provide
a principled way to deal with HRI scenario.  The framework is also
suitable for the development and experiment of task planners and
interaction schemes that explicitly consider human abilities and
preferences.

@@

We are planning to extend the scheme by taking into accoount situation where
divergent belief may appear between the human and the robot. Preliminary
results have been pusblished in~\cite{warnier2012when}.

There is also extensive work to be done in order to refine the notion of "good
shared plan" and "good/acceptable robt behaviour" in this context. There are
obviously large avenues for learning and adaptation in this context.

>>>>

\usepackage[nomargin, marginclue, footnote]{fixme}

\usepackage{hyperref}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\columnwidth]{grounding_robot.pdf}
\caption{Robot reasoning about HRI and anticipation of human activities.
  Sources of information are multi-modal dialogue, and monitoring of the
  environment and human activity. The robot must adapt on-line its behaviors
  by merging computed plans with reactive control. The robot explicitly reasons
  on the fact that it is (or not) observed by the human. Reasoning and planning
  take place at symbolic as well as geometric level and systematically take
  into account both agents beliefs and activities (from the point of view of
  the robot).}

@@

Connecting multiple independent software modules in one coherent robotic
architecture is not only a technical challenge, but also a design and
architectural challenge.

In particular, dealing with the rich semantics of natural interaction with
humans has been one of our major focus. To this end, we have been deploying
over the last years explicit knowledge representation and manipulation in the
deliberative layer of the robots as a {\it lingua franca} between our
components.

Figure~\ref{fig|archi} gives an overview of resulting
architecture~\cite{Alami2011}. An active knowledge base, conveniently thought
as a \emph{semantic blackboard}, connects most of the modules together: the
\emph{geometric reasoning} module (called {\sc Spark}) produces symbolic
assertions describing the dynamics of the robot's environment. These logical
statements are stored in the knowledge base, and reused by the language
processing module (called {\sc Dialogs}), the symbolic task planner (HATP) and
the execution controllers. The output of the language processing module, or the
activities started by the robot controller are likewise stored as symbolic
statements.

Incoming statements may in turn trigger events in the system, which
allow reactive behaviors. These mechanisms are illustrated in the coming
sections.

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{archi}
        \caption{Overview of the deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may in return
trigger physical behaviors.

@@

HATP originality resides in its ability to produce plans for the robot's
actions \emph{as well as for the other participants} (humans or robots): the
robot plans for itself, but also for its interactors, which effectively endows the robot
with new interaction capabilities: It can partially predict the behavior, and
thus monitor the engagement, of its human partner; the planner generates
synchronisation points (arrow going from the robot stream to the human stream on
Figure~\ref{plan_hatp1}) between the agents; the controller can also verbalize
the plans to explain its partner how a task may be shared~\cite{warnier2012when}.

@@

This allows us to write reactive robot controllers with a high level of
expressiveness: for instance, by subscribing to the event \stmt{human1 desires
?action, ?action type Look, ?action hasGoal myself}, we could trigger a
behavior when the human expresses (through dialogue, gestures...) that he
wants to look at the robot itself.

@@

The second one involves the SHARY execution controller and the HATP symbolic
task planner besides ORO and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. behaviors and motions
are fully planned.

@@

Cognitive control architectures for robotics is a research field in its own right.
Our originality of our contribution relates to the interleaving of a
semantic-aware architecture designed for autonomy with human-aware decisional
components. To our knowledge, no other robotic architecture covers in depth
those two perspectives.

The Ke Jia project~\cite{Chen2010} is maybe one of the closest approach. It
integrates on a mobile platform a knowledge representation language with
natural language processing, task planning and motion planning. Default and
non-monotonic reasoning has been especially researched within the Ke Jia
project for symbolic task planing~\cite{Ji2011} and underspecified natural
language processing.

The Ke Jia robot has been demonstrated in several tasks involving human-robot
interaction with natural language. These tasks include a task with multiple
\emph{pick \& carry} that are globally optimised, naive physics reasoning via
taught rules or more complex scenarios with the robot delivering drinks, taking
into account changing and mutually exclusive preferences of users.

Also notable, Beetz et al.~\cite{Beetz2010} proposes a cognitive architecture
called \textsc{CRAM} (Cognitive Robot Abstract Machine) that integrates
\textsc{KnowRob}~\cite{Tenorth2009a}, a knowledge processing framework based on
Prolog. Its underlying storage is based on an OWL ontology, derived from
\textsc{OpenCyc}. \textsc{CRAM} and \textsc{KnowRob} have been demonstrated on
several real-world scenarios, where natural language recipes extracted from the
Internet had to be translated into plans and executed in a kitchen environment,
perceived and rebuilt on-line by the robots. This architecture has however not
been deployed in scenarios involving human-robot interactions.

@@

In our implementation, perspective taking is tightly connected to the symbolic
knowledge models, and since our knowledge base allows for storage of one
knowledge model per agent, we have been able to endow the robot with a simple
theory of mind (as explained in section~\ref{sect|tom}): we explicitly model
what the robot knows about its partners in a symbolic way. This knowledge is
then re-used in different places, to correctly interpret what the human says,
or to plan tasks that are actually doable for the human.

The cognitive model that the robot builds for the agent it interacts with is
still simple and mostly focused on geometric features (who sees what? What are
our relative positions? etc.). Extending this knowledge with more subtle
perceptions (emotional state for instance) remains to be done.

We should also mention that modelling of human behaviors also appear in our
architecture at the level of task and motion planning with a set of
\emph{social rules}, including proxemics, action legibility or time
optimization.

@@

@@

Interaction with humans implies the ability to deal semantics: semantics of
verbal interaction, semantics of gestures, etc.  As a consequence, it also
implies to deal with semantic disambiguation.

@@

We have presented in this article a decisional framework for human-robot
interactive task achievement that is aimed to allow the robot not only to
accomplish its tasks but also to produce behaviors that support its engagement
vis-a-vis its human partner and to interpret human behaviors and intentions.
Together and in coherence with this framework, we have developed and
experimented various task planners and interaction schemes that allow the robot
to select and perform its tasks while taking into account explicitly the human
abilities as well as the constraints imposed by the presence of humans, their
needs and preferences.

@@

One direction that we would like to further investigate is how to account for
situations where divergent beliefs appear between the human and the robot.
Preliminary results have been published in~\cite{warnier2012when}.

There is also extensive work to be done in order to refine the notion of ``good
shared plan" and ``good/acceptable robot behavior" in this context. There are
obviously large avenues for learning and adaptation in this context.

Another broader direction to head deals with \emph{contexts representation}.
Contexts are currently often limited to the current spatial and temporal
situation. Some projects model offer the possibility to jump in the past or to
switch to another agent's perspective, but in current approaches, selecting a
context always basically consists in retrieving a set of beliefs corresponding
to a situation, and temporarily replacing the current beliefs by those other
ones. This misses the fact that at a given moment, not one but many contexts
co-exist at different scales. We do not want to retrieve one monolithic set of
beliefs, but instead carefully craft a context from several \emph{atomic}
contexts. Techniques for representation of overlapping pools of knowledge
largely remain to be developed, as well as efficient algorithms to retrieve (or
discard) such context-related pools of knowledge. This is a challenge not only for robotics, but more generally for artificial intelligence.

The ability to explicitly manage contexts and context switches would endow the
robot with a cognitive capability similar to what is known as
\emph{context-dependent memory} in cognitive psychology. This is also related
to Tulving's \emph{autonoetic consciousness}~\cite{Tulving1985a}: the ability
to reflect upon its own past or future experiences.

Much remain to be done to this regard, starting with a formal analysis of what
are the relevant contexts for our robots.

Human-Robot Interaction is and is going to remain a challenging field for
artificial intelligence. We hope that this contribution helps with clarifying
some of these challenges.
